Attempting to Predict Heart Disease using a Classical Decision Tree Machine Learning Method

### 2022-12-09 Manuscript/paper.tex
fixed a few more errors

In conclusion, the error rates of all three models hovered around 50\%. This is not a good result. Since the first error rate was so large in the first model with all of the variables I was led to try different models to see if they would improve the error rate. While each error rate is not identical I would argue that the differences between them are insignificant. I wanted to create a model without the cholesterol values of 0 mg/dL because if the values reported are truly the serum levels of cholesterol it would be impossible if not extremely unhealthy for these to be 0 mg/dL. Originally I thought the creator of the data frame might have switched individuals who maybe didn\textsc{\char13}t have cholesterol levels to report from NA to 0 just so the usability and number of NA values was 0. Additionally after getting 2 bad results I decided to reasses and reflect, go through my notes and think about what could possibly make this model better. I found that these models work best with variables that are either numeric where a cut off point can be determined or strictly binary variables. This led me to take out the chest pain type because it had 4 buckets instead of 2. I also took out the ST slope variable because it seemed redundant and it also had 3 buckets of classification instead of 2. This model technically ended up having the smallest error rate but it still was around 50\%. \par 
Limitations in this paper mainly stem from the data source itself. Since there are over 100 individuals with claimed blood serum cholesterol levels of 0 mg/dL when this is not medically possible I think that data point may have been misrepresented. Additionally even though there were a significant amount of variables provided I think the addition of more variables like weight, BMI, exercise frequency, or a quantitative measure of healthy lifestyle might make better less error prone models. I also think the method chosen to create the models was limiting. Since all of the models have a 50\% error rate I can conclude and reconfirm that classical decision trees are not be the best machine learning model for this data. At the time of writing the proposal for this project and then when developing the rough draft and doing the actual coding and model creation for this project I did not have the knowledge of many other methods of machine learning for binary variables at the time. As the semester has continued and my knowledge from other classes has expanded the breadth of statisitical processes that I know, it would be incredibly beneficial to delve into models such as logistic regression, linear discriminant analysis, quadratic discriminant analysis, and deep learning models such as feed foward networks. Finally while this data set was large and quite comprehensive with more time, money, and resources more individuals and variables could be added to the data set. \par
Finally and most importantly, this study as it stands in its current state proves something very important: computers won\textsc{\char13}t be overtaking a doctor\textsc{\char13}s role anytime soon. The grueling years of training and studying still prevails over a tiny microchip in a motherboard. The results of this study however should not discourage the motivation of trying to come up with statistical models for medical outcomes. This paper should merely mark a moment in time. In the future as statistical knowledge increases, as more patient data is released, and as computers get better, faster, and smarter, someday a computer may help a doctor in diagnosis of heart disease.

>>>>

In conclusion, the error rates of all three models hovered around 50\%. This is not a good result. Since the first error rate was so large in the first model with all of the variables I was led to try different models to see if they would improve the error rate. While each error rate is not identical I would argue that the differences between them are insignificant. I wanted to create a model without the cholesterol values of 0 mg/dL because if the values reported are truly the serum levels of cholesterol it would be impossible if not extremely unhealthy for these to be 0 mg/dL. Originally I thought the creator of the data frame might have switched individuals who maybe didn\textsc{\char13}t have cholesterol levels to report from NA to 0 just so the usability and number of NA values was 0. Additionally after getting 2 bad results I decided to reasses and reflect, go through my notes and think about what could possibly make this model better. I found that these models work best with variables that are either numeric where a cut off point can be determined or strictly binary variables. This led me to take out the chest pain type because it had 4 buckets instead of 2. I also took out the ST slope variable because it seemed redundant and it also had 3 buckets of classification instead of 2. This model technically ended up having the smallest error rate but it still was around 50\%. \par 
Limitations in this paper mainly stem from the data source itself. Since there are over 100 individuals with claimed blood serum cholesterol levels of 0 mg/dL when this is not medically possible I think that data point may have been misrepresented. Additionally even though there were a significant amount of variables provided I think the addition of more variables like weight, BMI, exercise frequency, or a quantitative measure of healthy lifestyle might make better less error prone models. I also think the method chosen to create the models was limiting. Since all of the models have a 50\% error rate I can conclude and reconfirm that classical decision trees are not be the best machine learning model for this data. At the time of writing the proposal for this project and then when developing the rough draft and doing the actual coding and model creation for this project I did not have the knowledge of many other methods of machine learning for binary variables at the time. As the semester has continued and my knowledge from other classes has expanded the breadth of statistical processes that I know, it would be incredibly beneficial to delve into models such as logistic regression, linear discriminant analysis, quadratic discriminant analysis, and deep learning models such as feed forward networks. Finally while this data set was large and quite comprehensive with more time, money, and resources more individuals and variables could be added to the data set. \par
Finally and most importantly, this study as it stands in its current state proves something very important: computers won\textsc{\char13}t be overtaking a doctor\textsc{\char13}s role anytime soon. The grueling years of training and studying still prevails over a tiny microchip in a motherboard. The results of this study however should not discourage the motivation of trying to come up with statistical models for medical outcomes. This paper should merely mark a moment in time. In the future as statistical knowledge increases, as more patient data is released, and as computers get better, faster, and smarter, someday a computer may help a doctor in diagnosis of heart disease.
